\bf Estimating the number of clusters using cross-validation

### 2019-01-11 JCGS-R2/response/response.tex

\section*{Response to Editor}

\begin{itemize}

\item \textit{Fix the figure captions.}

\textbf{Response:} Thank you for the guidance here. Per your instruction, all
figure captions now include (1)~what is the plot about, (2)~specific details
of plot, and (3)~the most important thing that the reader should learn.

@@

(No further comments.)

\section*{Response to Reviewer 2}

\begin{itemize}

\item \textit{Bottom of page 4. Clarify meaning of ``fold''.}

\textbf{Response:} Added the following:

``For each value $r$ in the range $1, \dotsc, K$, we define the $r$th `fold'
as the matrix gotten by permuting the rows of the original data matrix to get
$X$, a matrix with the $r$th test subset, the rows $i$ with $r_i = r$ as
its trailing rows.''

### 2019-01-11 JCGS-R2/cvclust.tex
add R package citations

We confirm the result of Proposition~\ref{prop:single-2d} with a simulation.
We perform $10$ replicates.  In each replicate, we generate $20000$
observations from a mean-zero bivariate normal distribution with unit marginal
variances and correlation $\rho$.  We perform a single $2 \times 2$ fold of
Gabriel cross-validation and report the cross-validation mean squared error
for the number of clusters $k$ ranging from $1$ to $5$.
Figure~\ref{fig:nullcorr-equal} shows the cross-validation errors for all $10$
replicates.  The simulation demonstrates that in the Gabriel cross-validation
criterion chooses the correct answer $k = 1$ whenever $\rho < 0.5$; the
criterion chooses $k \geq 2$ clusters whenever $|\rho| > 0.5$.

@@

In this section, we perform a set of simulations to evaluate the performance
of our proposed method and the associated correlation correction described in
Section~\ref{sec:corr-correct}.  We compare our method with a basket of
competing methods including the Gap statistic
\citep{tibshirani2001estimating}, Gaussian mixture model-based clustering
\citep{fraley2002model}, the CH-index \citep{calinski1974dendrite}, Hartigan's
statistic \citep{hartigan1975clustering}, the Jump method
\citep{sugar2003finding}, Prediction strength \citep{tibshirani2005cluster},
Slope method \citep{fujita2014non} and Bootstrap stability \citep{fang2012selection}. 
We use the default parameter settings for all competing methods.  For Gabriel cross-validation,
we perform $2$-fold cross-validation on the columns~($p=q$) and $5$-fold
cross-validation on the rows~($m=n/4$).  We also compare with Wold
cross-validation, which we describe in Appendix~\ref{sec:wold-cv}.

>>>>

We confirm the result of Proposition~\ref{prop:single-2d} with a simulation.
Here and throughout the text, our simulations use
R~software environment~\citep{rproject}, with most figures generated
with the support of the \texttt{ggplot2} library~\citep{ggplot2}.
We perform $10$ replicates.  In each replicate, we generate $20000$
observations from a mean-zero bivariate normal distribution with unit marginal
variances and correlation $\rho$.  We perform a single $2 \times 2$ fold of
Gabriel cross-validation and report the cross-validation mean squared error
for the number of clusters $k$ ranging from $1$ to $5$.
Figure~\ref{fig:nullcorr-equal} shows the cross-validation errors for all $10$
replicates.  The simulation demonstrates that in the Gabriel cross-validation
criterion chooses the correct answer $k = 1$ whenever $\rho < 0.5$; the
criterion chooses $k \geq 2$ clusters whenever $|\rho| > 0.5$.

@@

In this section, we perform a set of simulations to evaluate the performance
of our proposed method and the associated correlation correction described in
Section~\ref{sec:corr-correct}.  We compare our method with a basket of
competing methods including the Gap statistic
\citep{tibshirani2001estimating}, Gaussian mixture model-based clustering
\citep{fraley2002model}, the CH-index \citep{calinski1974dendrite}, Hartigan's
statistic \citep{hartigan1975clustering}, the Jump method
\citep{sugar2003finding}, Prediction strength \citep{tibshirani2005cluster},
Slope method \citep{fujita2014non} and Bootstrap stability \citep{fang2012selection}. 
We use the default parameter settings for all competing methods, implented
in the following R~packages:
\texttt{cluster} \citep{cluster};
\texttt{cstab} \citep{cstab};
\texttt{e1071} \citep{e1071};
\texttt{fpc} \citep{fpc};
\texttt{MASS} \citep{MASS};
\texttt{mclust} \citep{mclust};
and
\texttt{NbClust} \citep{NbClust}.
For Gabriel cross-validation,
we perform $2$-fold cross-validation on the columns~($p=q$) and $5$-fold
cross-validation on the rows~($m=n/4$).  We also compare with Wold
cross-validation, which we describe in Appendix~\ref{sec:wold-cv}.
